---
title: "Class Activity 411"
author: "Rickey Huang"
date: "4/11/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data <- read.csv(file = "houses.csv")
head(data)
```

\section{Question 1}
\subsection{Part a}

*Fit a model Price ~ Living.Area*

```{r}
lm1 <- lm(Price~Living.Area, data = data)
```

\subsection{Part b}

*Plot ehat vs. fitted*

```{r,fig.asp=0.6}
plot(lm1$residuals,lm1$fitted)
```

\subsection{Part c}

*Make a residual histogram*

```{r,fig.asp=0.6}
hist(lm1$residuals)
```

\subsection{Part d}

*Make a QQ-plot*

```{r,fig.asp=0.55}
e1 <- lm1$residuals
e1 <- sort(e1)
n <- dim(model.matrix(lm1))[1]
Q <- qnorm(seq(1:n)/(n+1))
sigmahat <- sqrt(sum(e1^2)/(n-2))
z <- e1/sigmahat
plot(Q,z)
abline(0,1)
```

\subsection{Part e}

*Can you see any bad problem individuals?*



\subsection{Part f}

*What happens if you remove them?*

```{r}

```

\section{Question 2}
\subsection{Part a}

*Create a vector of internally studentized residuals*

```{r}

```

\subsection{Part b}

*Creaate a vector of externally studentized residuals*

\subsection{Part c}

*Create a vector of Cook's Distance*

\subsection{Part d}

*Do the large values correspond to the data points you thought were bad from last time?*

\subsection{Part e}

*What happends if you remove them?*


