---
title: "Class Activity Week 3"
author: "Rickey Huang"
date: "1/26/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```

## 1. Randomly generate data on a line 
```{r}
x = runif(100, 5, 15)
y = rnorm(100,x+2*x^2, 50)
```

## Plot the data, fit a parabola
```{r}
ord=order(x)
xord=x[ord]
yord=y[ord]
plot(xord,yord)
lm1 <- lm(yord~xord + I(xord^2))
points(xord, lm1$fitted, type = "l",lwd=3, col = "orange")
```

## Extract fitted values for y, or our predictions based on the parabola

```{r}
fittedy <- lm1$fitted
fittedy
```

## Define the model matrix x

```{r}
X = cbind(1,x)
X
```

## compute hat matrix H and yhat

```{r}
H <- X%*%solve(t(X)%*%X)%*%t(X)
H
```

## Compare lm1$fitted and yhat

```{r}
yhat <- H%*%y
## Compare fittedy and yhat
yhat == fittedy
```

## Compare HX and X

```{r}
HX <- H%*%X
## Comparison
HX == X
```

## Comapre Hyhat and yhat
```{r}
Hyhat <- H%*%yhat
## Comparison
Hyhat == yhat
```

## Compare residual and matrix definition of ehat

```{r}
rehat <- y - yhat
I = diag(100)
ehat <- (I - H)%*%y
rehat == ehat
```

## Compute the Sum of Squared Estimated Errors

```{r}
SSE <- t(y)%*%(I-H)%*%y
SSE
```

## Compute the Sum of Squres for Regression

```{r}
SSR <- (yhat - )
```

## Compute the Sum of Squares Total
